---
title: Habenula Atlas Visium
output:
  github_document:
    html_preview: true
  html_document:
    toc: true
    toc_flot: true
    includes:
      in_header: header.html
      after_body: footer.html
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r 'setup', include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "img/",
    out.width = "100%"
)
```

**Zenodo DOI TODO**

## Overview

Welcome to the Habenula Atlas Visium and Visium HD repository! This repository and its [multiome companion](https://github.com/LieberInstitute/Hb_multiome) contain all data analysis code used in our manuscript where we produced an atlas of the human habenula. We generated spatial transcriptomics data using the 10x Genomics Visium and Visium HD platforms, as well as single-nucleus multiome data (snRNA-seq + snATAC-seq) from a total of 13 neurotypical control donors.

## Study Design

**TODO: fig 1 and caption**

## Interactive Websites

All of these interactive websites are powered by open source software, namely:

* 🔭 [`spatialLIBD`](https://doi.org/10.1186/s12864-022-08601-w)
* 👀 [`iSEE`](https://doi.org/10.12688%2Ff1000research.14966.1)
* 🧬 [`shiny`](https://doi.org/10.32614/CRAN.package.shiny)

We provide the following interactive websites, organized by dataset with software labeled by emojis:

* Visium (n = 3)
  - 🔭 [Habenula_Visium](https://interactive.libd.org/Habenula_Visium): [`spatialLIBD`](https://doi.org/10.1186/s12864-022-08601-w) website showing the spot-based Visium results.
* Visium HD (n = 5)
  - 🔭 [Habenula_Visium_HD](https://interactive.libd.org/Habenula_Visium_HD): [`spatialLIBD`](https://doi.org/10.1186/s12864-022-08601-w) website showing the cell-level Visium HD results.
  - 👀 [Habenula_Visium_HD_pb](https://interactive.libd.org/Habenula_Visium_HD_pb): [`iSEE`](https://doi.org/10.12688%2Ff1000research.14966.1) website showing the Visium HD results pseudobulked by cell type.
* Multiome (n = 10)
  - 🧬 [Habenula_multiome](https://interactive.libd.org/Habenula_multiome): [`shiny`](https://doi.org/10.32614/CRAN.package.shiny) app showing the snRNA-seq and snATAC-seq data, aggregated at various levels. Reduced dimension plots show cell-level data, trio results are shown at the metacell level, and differentially accessible regions (DARs) are explorable with the data pseudobulked by cell type.

### Local `spatialLIBD` apps
  
If you are interested in running the [`spatialLIBD`](https://doi.org/10.1186/s12864-022-08601-w) applications locally, you can do so thanks to the [`spatialLIBD::run_app()`](http://research.libd.org/spatialLIBD/reference/run_app.html), which you can also use with your own data as shown in our [vignette for publicly available datasets provided by 10x Genomics](http://bioconductor.org/packages/release/data/experiment/vignettes/spatialLIBD/inst/doc/TenX_data_download.html).

**TODO: add `run_app()` commands for the real data we're sharing**

## Contact

We value public questions, as they allow other users to learn from the answers. If you have any questions, please ask them at [LieberInstitute/Habenula_Visium/issues](https://github.com/LieberInstitute/Habenula_Visium/issues) or [LieberInstitute/Habenula_multiome/issues](https://github.com/LieberInstitute/Habenula_multiome/issues) and refrain from emailing us. Thank you again for your interest in our work!

## Citing our work

**TODO: our manuscript**

### Cite `spatialLIBD`

Below is the citation output from using `citation('spatialLIBD')` in R. Please run this yourself to check for any updates on how to cite __spatialLIBD__.

```{r 'citation_spatialLIBD'}
print(citation("spatialLIBD")[1], bibtex = TRUE)
```

Please note that the `spatialLIBD` was only made possible thanks to many other R and bioinformatics software authors, which are cited either in the vignettes and/or the paper(s) describing the package.

## Data Access

**TODO**

## Internal

* JHPCE locations:
  - `/dcs04/lieber/lcolladotor/Habenula_R01_LIBD4270/Habenula_Visium`
  - `/dcs04/lieber/lcolladotor/Habenula_R01_LIBD4270/Hb_multiome`
